A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-13 docs/icccn2009/paper.tex
Adding better quality EPS images and adjusting their sizes.
Adding 2 the header image in case we wanna add more details
about it.

git-svn-id: svn://localhost/netbeams/trunk@377 559edde2-a4d4-44e0-96fc-4035e40d7396
Incorporated Dale's feedback
git-svn-id: svn://localhost/netbeams/trunk@379 559edde2-a4d4-44e0-96fc-4035e40d7396

The work described in this paper is a collaborative effort between the
Computer Science Department at San Francisco State University and the
Romberg Tiburon Center (RTC), a research institute focusing on the
understanding of complex marine and estuarine environments. The result
of this collaboration is NetBEAMS (Networked Bay Environmental
Assessment Monitoring System), an environmental sensor network
deployed in the San Francisco Bay. NetBEAMS provides means to program
and interrogate sensors at the remote field locations and to transmit
collected measurements to a database, where the data is rapidly
processed, archived, and made available to potential users in near
real-time. In the following we give an overview of the NetBEAMS
architecture and implementation.  Section \ref{SEC_BACKGROUND}
describes a typical environmental sensor used by the RTC. Section
\ref{SEC_DSP} describes our Data Sensor Platform (DSP) that allows us
to efficiently build up an end-to-end environmental sensor network.
Section \ref{SEC_NETBEAMS} describes NetBEAMS; a practical application
of the DSP. Section \ref{SEC_CONCLUSION} concludes this paper.

@@

Environmental researchers typically use standard, off-the-shelf
sensing devices for their purposes. These sensing devices are at the
opposite spectrum of miniature sensors such as SmartDust. One such
device is the YSI 6600EDS V2 Sonde \cite{Sonde01}. It is a water
quality monitoring device that gathers water quality data, and
operates in fresh, sea, or polluted water. Some of the measurements
that the YSI is capable of are conductivity, temperature, chloride,
ammonium, nitrate, turbidity, and chlorophyll.  It is under 55 cm in
length, 8.9 cm in diameter, and weighs approximately 3.18 kg. It
operates at temperatures between -5 to +50 deg C and at depths up to
200 meters.  It uses 8 C-size Alkaline Batteries or External 12 VDC.
Battery life can last up to 75 days depending on sensor configuration.
The Sonde's bulkhead contains multiple pin connectors to support
sensor probes for measuring parameters as shown in Table
\ref{TAB_SONDE_MEASUREMENTS}.

@@

\begin{figure}[!htb]
\centering
\epsfig{file=dsp-message-body-ysi, width=8cm, height=7cm}
\caption{\label{FIG_DSP_YSI_DATA} YSI message format.}
\end{figure}

@@

All messages sent by the YSI Sonde DP need to be uploaded to a remote
database. Since the DSP itself can only handle communication within
its local DPs and DCs, we have created a matching pair of DC and DP
that can handle remote communication (components (2) and (3) in Figure
\ref{FIG_NETBEAMS}). As explained earlier, we make use of the cellular
phone network to upload sensor data to a remote backend server. While
the cellular phone network has sufficient coverage for more
environmental sensor networks, it is not cost effective to keep a data
link permanently open. The remote communiction components can be
configured to upload buffered sensor data only in certain time
intervals.Messages are transported via HTTP \cite{RFC2068}. The
matching DP/DC pair for remote communication has the following
properties:

>>>>

The work described in this paper is a collaborative effort between the
Computer Science Department at San Francisco State University and the
Romberg Tiburon Center (RTC), a research institute focusing on the
understanding of complex marine and estuarine environments. The result
of this collaboration is NetBEAMS (Networked Bay Environmental
Assessment Monitoring System), an environmental sensor network
deployed in the San Francisco Bay. NetBEAMS provides means to program
and interrogate sensors at the remote field locations and to transmit
collected measurements to a database, where the data is rapidly
processed, archived, and made available to potential users in near
real-time. In the following sections we give an overview of the NetBEAMS
architecture and implementation.  Section \ref{SEC_BACKGROUND}
describes a typical environmental sensor used by the RTC. Section
\ref{SEC_DSP} describes our Data Sensor Platform (DSP) that allows us
to efficiently build up an end-to-end environmental sensor network.
Section \ref{SEC_NETBEAMS} describes NetBEAMS; a practical application
of the DSP. Section \ref{SEC_CONCLUSION} concludes this paper.

@@

Environmental researchers typically use standard, off-the-shelf
sensing devices for their purposes. These sensing devices are at the
opposite end of the size spectrum of miniature sensors such as SmartDust. One such
device is the YSI 6600EDS V2 Sonde \cite{Sonde01}, a water
quality monitoring device that gathers water quality data, and
operates in fresh, sea, or polluted water. Some of the measurements
that the YSI is capable of are conductivity, temperature, chloride,
ammonium, nitrate, turbidity, and chlorophyll.  It is under 55 cm in
length, 8.9 cm in diameter, and weighs approximately 3.18 kg. It
operates at temperatures between -5 to +50 deg C and at depths up to
200 meters.  It uses 8 C-size Alkaline Batteries or External 12 VDC.
Battery life can last up to 75 days depending on sensor configuration.
The Sonde's bulkhead contains multiple pin connectors to support
sensor probes for measuring parameters as shown in Table
\ref{TAB_SONDE_MEASUREMENTS}.

@@

\begin{figure}[!htb]
\centering
\epsfig{file=dsp-message-body-ysi, width=9cm}
\caption{\label{FIG_DSP_YSI_DATA} YSI message format.}
\end{figure}

@@

All messages sent by the YSI Sonde DP need to be uploaded to a remote
database. Since the DSP itself can only handle communication within
its local DPs and DCs, we have created a matching pair of DC and DP
that can handle remote communication (components (2) and (3) in Figure
\ref{FIG_NETBEAMS}). As explained earlier, we make use of the cellular
phone network to upload sensor data to a remote backend server. While
the cellular phone network has sufficient coverage for more
environmental sensor networks, it is not cost effective to keep a data
link permanently open. The remote communiction components can be
configured to upload buffered sensor data only in certain time
intervals. Messages are transported via HTTP \cite{RFC2068}. The
matching DP/DC pair for remote communication has the following
properties:
